---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(rdoom)
```

# rdoom

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
<!-- badges: end -->


`{rdoom}` is a playable version of [Doom](https://en.wikipedia.org/wiki/Doom_(1993_video_game)) 
for R.   This package includes the demo game, but it should also work with the
game `.wad` file for DoomI, DoomII and related games.

This is a technical demonstration for the rendering speed and interactivity of [tigerfb](https://github.com/coolbutuseless/tigerfb).

`{tigerfb}` is a cross-platform framebuffer for R which enables realtime, interactive 
graphical displays in R with mouse & keyboard input.

It uses:

* [`doomgeneric`](https://github.com/ozkl/doomgeneric) as the game engine. A fork
  of the original Doom source code.
* [`{tigerfb}`](https://github.com/coolbutuseless/tigerfb) as the framebuffer for R
* [`{nara}`](https://cran.r-project.org/package=nara) for fast native raster image support
* [`{audio}`](https://cran.r-project.org/package=audio) for playing sound effects


This code should run on macos, linux and windows platforms (but has only been
extensively tested on macos).

Suggested way to run this on all platforms: 

```{r eval = FALSE}
install.packages(c('nara'))                # Backend render surface
install.packages(c('audio'))               # Sound support
pak::pkg_install('coolbutuseless/tigerfb') # Framebuffer
pak::pkg_install('coolbutuseless/rdoom')   # This doom engine

# Run with the demo WAD file
rdoom::doom()
```




## Quirks

#### Mouse control

The mouse control is a bit awkward.  This doom engine is expecting that the mouse
movement delta between frames is always available.  However the framebuffer
library being used by `{tigerfb}` only supports reading the mouse cursor
position on the screen.  This means that when the mouse hits the edge of the
screen, the movement input will fail.

Fix: the best interim fix I implemented is that mouse movement in the bottom 
fraction of the screen does not act as input to the game.  This means you 
can re-centre the mouse by moving it low-down on the screen.


### Cheat Codes

Most cheat codes won't work directly because of the way keyboard handling is
coordinated between the doom engine and the frame buffer provided by `{tigerfb}`

To enter a cheat code, such as `iddqd`, you need to do the following:

1. Press `F9` to temporarily disable control via the keyboard
2. Type `i`, `d`, `d`, `q`, `d`
3. Press `F10` to re-enable the keyboard controls


### Audio

For sound effects, `{rdoom}` exploits the fact that the `{audio}` package 
can play sound asynchronously.

The downside of the `{audio}` package is that it only does basic additive 
sound mixing, so when multiple sounds play at the same time, the combined 
magnitude exceeds the limit of the audio device and *clipping* makes
it all sound awful.

To workaround this, all audio samples have been scaled to a lower volume, and
there is a minimum 0.1 second window between sound triggers.

This means that sound may be softer than expected, and when many sounds are
triggered to play simultaneously some of them will just go missing.














